# Assignment 4: LAGOS Spacial Analysis


```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
setwd("/Users/mjcarlson/Desktop/ESS580 Environmental Data Science/ESS580Bookdown")
```


## Part 1 LAGOS Analysis


### Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
#LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus


```



### Convert to spatial data
```{r}
#Look at the column names
names(lake_centers)

#Look at the structure
str(lake_centers)

#View the full dataset
View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



## In-Class work


### 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
#mapview(states)
Icombined <- states %>%
  filter(name %in% c('Illinois', 'Iowa')) %>%
  st_transform(2163)

I_lakes <- spatial_lakes[Icombined,]

#Iowa
Iowa <- states %>%
  filter(name == 'Iowa') %>%
  st_transform(2163)

#Illinois
Illinois <- states %>%
  filter(name == 'Illinois') %>%
  st_transform(2163)

#rbind states
IA_IL<-rbind(Iowa,Illinois)
mapview(IA_IL)

```



### 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa  combined? How does this compare to Minnesota?

```{r}
illinois_lakes <- spatial_lakes[Illinois,]
iowa_lakes <- spatial_lakes[Iowa,]

#rbind lake data
IA_IL2<-rbind(iowa_lakes, illinois_lakes)

#count
str(IA_IL2$lagoslakeid)
str(minnesota_lakes$lagoslakeid)

```

There are about twice as many lakes in Minnesota, 29038, as there are in Iowa and Illinois combined, 16466.


### 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
ggplot(iowa_lakes, aes(x=lake_area_ha))+
  geom_histogram()+
  scale_x_log10()+
  labs(title = "Iowa",
    x="log(lake area(ha))")

ggplot(minnesota_lakes, aes(x=lake_area_ha))+
  geom_histogram()+
  scale_x_log10()+
  labs(title = "Minnesota",
    x="log(lake area (ha))")

```

There are more larger lakes in Minnesota than in Iowa where the lakes are much smaller.


### 4) Make an interactive plot of lakes in Iowa and Illinois and color them  by lake area in hectares

```{r}
I_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')

```


### 5) What other data sources might we use to understand how reservoirs and  natural lakes vary in size in these three states? 

Using information such as the perimeter or the lake and lake depth would give another source of data to compare the sizes of the lakes.


# Part 2 LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')


```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))%>%
  #Take the log base 10 of the mean_secchi
  mutate(log10_mean_secchi = log10(mean_secchi))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')
```


## Class work

### 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for  sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r, warning=F}
#Your code here
ggplot(chla_secchi_200, aes(y = chla, x = secchi))+
  geom_point(size = 1, shape = 1)+
  scale_x_log10()+
  scale_y_log10()
```


### Why might this be the case? 

When there is more chlorophyll in the water it is more turbid and therefore harder to see the Secchi disk.

### 2) What states have the most data? 

### 2a) First you will need to make a lagos spatial dataset that has the total  number of counts per site.

```{r, warning=FALSE}
## Your code here
site_counts<-nutr%>%
  group_by(lagoslakeid)%>%
  summarize(count = n())

spatial_counts<-inner_join(spatial_lakes, site_counts%>%
                             distinct(lagoslakeid,.keep_all = T),
                           by = "lagoslakeid")

```


### 2b) Second, you will need to join this point dataset to the us_boundaries  data. 

```{r, warning=F}
## Your code here
states_lagos<-lagos$state

states_lakes <- inner_join(spatial_counts, states_lagos, by="state_zoneid")
```


### 2c) Then you will want to group by state and sum all the observations in  that state and arrange that data from most to least toatl observations per state. 

```{r}
## Your code here. 
sum_states<-states_lakes%>%
  as.data.frame()%>%
  select(-geometry)%>%
  group_by(state_name)%>%
  summarize(sum_counts = sum(count))%>%
  arrange(desc(sum_counts))

head(sum_states)
```

Minnesota has the most data followed by Wisconsin and Michigan

### 3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200  observations?

```{r}
## Your code here
mean_spatial%>%
  arrange(-mean_secchi) %>%
  mapview(.,zcol = 'mean_secchi')

mapview(mean_spatial,zcol='log10_mean_secchi')

```

There is not a spatial pattern in Secchi disk depths, there are some clusters of locations where there are many lakes with more than 200 observations but this has more to do with lake density than Secchi disk depths.


