--- 
title: "ESS 580 Bookdown"
author: "Madeline Carlson"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
--- 
title: "ESS 580 Bookdown"
author: "Madeline Carlson"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Introduction

This is a book of all the assignments done in Introduction to Environmental 
Data Science (ESS 580A-7).


<!--chapter:end:index.Rmd-->


# Assignment 1: R Markdown Examples

Placeholder


## Methods
### SiteDescription
### Data Acquisition and Plotting tests
### Data Download
### Static Data Plotter
### Interactive Data Plotter
## Assignment. 
## DyGraph example. 
## Poudre Paragraph

<!--chapter:end:02-rmarkdown_examples.Rmd-->


# Assignment 2: Fire Data Wrangle

Placeholder


## Question 1) 
## Question 2 
## Q3
## Question 4
## Question 5) 

<!--chapter:end:03-fire_data_wrangle.Rmd-->

# Assignment 3: Snow Function Iteration


```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)
setwd("/Users/mjcarlson/Desktop/ESS580 Environmental Data Science/ESS580Bookdown")
```


### Simple web scraping

R can read html using either rvest, xml, or xml2 packages. Here we are going to navigate to the Center for Snow and Avalance Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data we want to programatically download for three sites. We don't know much about these sites, but they contain incredibly rich snow, temperature, and precip data. 


## Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')
links
```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r}
#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('/Users/mjcarlson/Desktop/ESS580 Environmental Data Science/ESS580Bookdown/data_snow/',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


3. Write a custom function to read in the data and append a site column to the data. 

```{r}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

```


```{r}
meteorological_reader<-function(file_names){
  name=str_split_fixed(file_names,'/',2)[,2]
  name2=str_split_fixed(file_names,'/',4)[,2]
  df<-read.delim(file_names, header = F, sep = "", col.names = headers, skip = 4)%>%
    select(1:14)%>%
    mutate(site=name2)
}

```


4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.
 
```{r}

fulldata<-map_dfr(file_names, meteorological_reader)

unique(fulldata$site)

```



5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r}
annual_mean<-fulldata%>%
  group_by(site, year)%>%
  summarize(mean_air_temp=mean(air.temp..K.))

ggplot(annual_mean, aes(x=year, y=mean_air_temp, color=site)) + 
  geom_line(size=2)+
  labs(title="Annual mean air temperature",
       x="Year",
       y= "Temperature (K)")+
  theme(legend.position = c(.7, .25))
  
```

2003 is much colder than the other years which is suspicious. When looking at the data we can see that the data starts in November so it is not a complete year of data and only includes two colder months.

```{r}
annual_mean<-fulldata%>%
  group_by(site, year)%>%
  filter(year>2003)%>%
  summarize(mean_air_temp=mean(air.temp..K.))

ggplot(annual_mean, aes(x=year, y=mean_air_temp, color=site)) + 
  geom_line(size=2)+
  labs(title="Annual mean air temperature",
       x="Year",
       y= "Temperature (K)")+
  theme(legend.position = c(.7, .25))
```


6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Swamp Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

```{r, warning=FALSE,message=FALSE}

#monthly_mean<-fulldata%>%
  group_by(month, year, site)%>%
  summarize(monthly_air_temp = mean(air.temp..K.))

ggplot(monthly_mean, aes(x = month, y = monthly_air_temp, color = site)) +
  facet_wrap(~year) +
  geom_line()

monthplot<-function(fulldata, year){
  monthlytemp<-fulldata%>%
    group_by(month, year, site)%>%
    summarize(monthly_air_temp = mean(air.temp..K.))%>%
    filter(yr == year)
  
  plots<-ggplot(monthlytemp, aes(x = month, y = monthly_air_temp, color = site))+
    geom_line(size=2)+
    labs(title = monthlytemp$year,
         x="Month",
         y="Temperature (K)")
  
  print(plots)
}

years<-c(2005, 2006, 2007, 2008, 2009, 2010)

for (yr in years){
  monthplot(fulldata, year)
}
```


<!--chapter:end:04-snow_function_iteration.Rmd-->


# Assignment 4: Part 1 LAGOS Spacial Analysis

Placeholder


## LAGOS Analysis
### Loading in data, first download and then specifically grab the locus (or site lat longs)
### Convert to spatial data
### Subset to only Minnesota
## In-Class work
### 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)
### 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa  combined? How does this compare to Minnesota?
### 3) What is the distribution of lake size in Iowa vs. Minnesota?
### 4) Make an interactive plot of lakes in Iowa and Illinois and color them  by lake area in hectares
### 5) What other data sources might we use to understand how reservoirs and  natural lakes vary in size in these three states? 

<!--chapter:end:05-LAGOS_spacial_analysis.Rmd-->


# Assignment 4: Part 2 LAGOS Spacial Analysis

Placeholder


### Loading in data, first download and then specifically grab the locus (or site lat longs)
### Subset columns nutr to only keep key info that we want
### Keep sites with at least 200 observations 
### Join water quality data to spatial data
### Mean Chl_a map
## Class work
### 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for  sites with at least 200 observations?
### Why might this be the case? 
### 2) What states have the most data? 
### 2a) First you will need to make a lagos spatial dataset that has the total  number of counts per site.
### 2b) Second, you will need to join this point dataset to the us_boundaries  data. 
### 2c) Then you will want to group by state and sum all the observations in  that state and arrange that data from most to least toatl observations per state. 
### 3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200  observations?

<!--chapter:end:06-LAGOS_spacial_analysis.Rmd-->


# Assignment 5: Weather Corn Regression

Placeholder


## Weather Data Analysis
### Load the PRISM daily maximum temperatures
## Temperature trends
### Summer temperature trends: Winneshiek County
### Winter Temperatures - Winneshiek County
### Multiple regression -- Quadratic time trend
### Download NASS corn yield data
## Assignment
### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?
### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 
### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.
### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.
### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.
### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

<!--chapter:end:07-weather_corn_regression.Rmd-->

